---
title: "Exploring U.S. Supreme Court Decisions"
author: "Gustavo Arruda"
output: github_document
---

## Get the data

```{r}
# load useful packages
library(tidyverse)
library(lubridate)

# load data
case_data <- read_csv("data/scdb-case.csv")
vote_data <- read_csv("data/scdb-vote.csv")
```

## Recode variables as you find necessary

```{r}

```

## What percentage of cases in each term are decided by a one-vote margin (i.e. 5-4, 4-3, etc.)

```{r}

case_data %>%
  mutate(difference = majVotes - minVotes) %>%
  group_by(term) %>%
  summarize(one_vote_margin = sum(difference == "1") / n()) %>%
  ggplot() +
  geom_line(aes(x = term, y = one_vote_margin, color = one_vote_margin)) +
  labs(title = "Percentage of one-vote margin decisions per term", x = "Term", y = "Percentage", color = "Percentage")
  
```

## For justices [currently serving on the Supreme Court](https://www.supremecourt.gov/about/biographies.aspx), how often have they voted in the conservative direction in cases involving criminal procedure, civil rights, economic activity, and federal taxation?

Organize the resulting graph by justice in descending order of seniority. (Note that the chief justice is always considered the most senior member of the court, regardless of appointment date.)

```{r}

combined_data <- inner_join(x = vote_data, y = case_data) %>% #connecting issueArea to vote data
  filter(justice == 108 | (justice >= 110 & justice <= 116) & (issueArea == 1 | issueArea == 2 | issueArea == 8 | issueArea == 12)) %>%
  group_by(justiceName) %>%
  mutate(justiceName = factor(justiceName, levels = c("JGRoberts", "CThomas", "SGBreyer",                                                    "SAAlito", "SSotomayor", "EKagan","NMGorsuch","BMKavanaugh"))) %>%
  summarize(conservative_freq = sum(direction == "1", na.rm = TRUE) / n())

ggplot(combined_data) +
  geom_col(aes(x = justiceName, y = conservative_freq, fill = conservative_freq)) +
  labs(title = "Ratio of Conservative Votes per Justice", x = "Justice Name", y = "Ratio of Votes", fill = "Conservative Ratio") +
  coord_flip()
  

```

## In each term, how many of the term's published decisions (decided after oral arguments) were announced in a given month?

```{r}

dated_case_data <- case_data %>%
  mutate(dateDecision = mdy(dateDecision)) %>%
  mutate(monthDecision = month(dateDecision)) %>%
  mutate(monthDecision = factor(monthDecision,
    levels = c(10:12, 1:9),
    labels = c(month.abb[10:12], month.abb[1:9]))) %>%
  group_by(term, monthDecision) %>%
  count()

  ggplot(data=dated_case_data) +
  geom_line(aes(x = term, y = n, color = n)) +
  labs(title = "Decisions per month per year", x = "Year", y = "Decisions per Month", color = "Count") +
  facet_wrap(~monthDecision)
  

```

## Which justices are most likely to agree with with the Court's declaration that an act of Congress, a state or territorial law, or a municipal ordinance is unconstitutional? Identify all cases where the Court declared something unconstitutional and determine the ten justices who most and least frequently agreed with this outcome as a percentage of all votes cast by the justice in these cases. Exclude any justice with fewer than 30 votes in cases where the Court's outcome declares something unconstitutional.

```{r}

unCon_data <- case_data %>%
  inner_join(vote_data) %>% #coordinate individual justices with constitutionality ruling
  group_by(justiceName) %>% #using individual justices as frame of analysis
  filter(declarationUncon != 1 & n() > 30) %>% #filtering out unwanted cases
  summarize(agree_unCon = sum(vote == 1, na.rm = TRUE) / n()) #creating a "proportion of agreement" variable, while avoiding NA messing up the analysis

most_agr <- slice_max(unCon_data, n = 10, agree_unCon) #slicing top 10 most agreeable
least_agr <- slice_min(unCon_data, n = 10, agree_unCon) #slicing top 10 least agreeable

ggplot(most_agr) +
  geom_col(aes(x = justiceName, y = agree_unCon, fill = agree_unCon)) +
  labs(title = "Unconstitutionality Agreement", x = "Justice Name", y = "Ratio of Agreements", fill = "Agreements") +
  coord_flip()

ggplot(least_agr) +
  geom_col(aes(x = justiceName, y = agree_unCon, fill = agree_unCon)) +
  labs(title = "Unconstitutionality Disagreement", x = "Justice Name", y = "Ratio of Agreements", fill = "Agreements") +
  coord_flip()
  
```

## In each term he served on the Court, in what percentage of cases was Justice Antonin Scalia in the majority?

```{r}

```

## Create a graph similar to above that adds a second component which compares the percentage for all cases versus non-unanimous cases (i.e. there was at least one dissenting vote)

```{r}

```

## In each term, what percentage of cases were decided in the conservative direction?

```{r}

```

## The Chief Justice is frequently seen as capable of influencing the ideological direction of the Court. Create a graph similar to the one above that also incorporates information on who was the Chief Justice during the term.

```{r}

```

## Session info

```{r, echo = TRUE}
devtools::session_info()
```


